---
title: "Bowel cancer analysis"
output: html_document
date: "2025-10-24"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#to make the renv intially
#renv::init()

#renv friendly loading of packages.
#dont use pacman for this as it may overwrite the version of the package used in renv
required_packages <- c("tidyverse", "here", "renv","Hmisc","gtsummary","brms","mice","posterior","marginaleffects","bayestestR")

missing <- required_packages[!sapply(required_packages, requireNamespace, quietly = TRUE)]
if (length(missing) > 0) {
  stop("Missing packages: ", paste(missing, collapse = ", "), 
       ". Please run renv::restore() to install them.")
}
#if a package isnt installed in renv::restore(), do this: 
#renv::install("tidyverse")

library(renv)
library(tidyverse)
library(here)
library(Hmisc)
library(gtsummary)
library(brms)
library(mice)
library(posterior)
library(marginaleffects)
library(bayestestR)

#function to test the convergence of each model in brm_multiple
convergence <- function(chains=4,imps=20,mod=mod){
  draws <- as_draws_array(mod)
  # Subset chains for each dataset
  draws_per_dat <- lapply(1:imps, \(i) {
    start_chain <- (i - 1) * chains + 1
    end_chain <- i * chains
    subset_draws(draws, chain = start_chain:end_chain)
  })
  summaries <- lapply(draws_per_dat, summarise_draws, default_convergence_measures())
  return(summaries)
}

```

# Analysis phase 1
## Read in the data
```{r}
bowel <- readRDS(here("Data","bowel.RDS"))

# bowel2 <- bowel %>% select(record_id,los)
# test <- full_join(bowel2,bo_p2_imp,by="record_id") %>%
#   group_by(record_id,.imp) %>%
#   mutate(summary_base = max(ifelse(time=="0",summary_score,NA),na.rm=T)) %>%
#   filter(time %in% c(1),.imp==1)

```

## Primary outcome - LOS
no one missing the primary outcome or covars
group, stoma, type of surgery, hospital
```{r}
#select vars for model and make the groups unblinded
bowel_mod <- bowel %>%
  dplyr::select(grp,redcap_data_access_group,stoma.factor,
                surgery_type.factor,los)%>%
  mutate(grp = ifelse(grp == "A","Treatment","Control"))

#run the brms model
los_mod <- brm(los ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
               data=bowel_mod,family=poisson,iter=10000,seed=46137112,
               prior = c(prior(normal(0,1000),class=b),
                         prior(student_t(3,0,100),class=Intercept)))

#test <- loo(los_mod)
#tests <- loo(los_mods) #worse with summary_base, p_loo 32.9 without, 37.3 with

#checking fit with this and rhat + ess. Want rhat < 1.05 and ess > 10000 for 95% CrI
#plot(los_mod)
#pp_check(los_mod, ndraws=100)

#rate ratio
losrr <- avg_comparisons(los_mod,variables=list(grp="pairwise"),
                         comparison = "ratio",ndraws=5000) %>%
  mutate(RR = paste0(sprintf("%.2f",round(estimate,2))," (", round(conf.low,2),", ",round(conf.high,2),")")) %>%
  dplyr::select(RR)
#p(sup)
losh <- hypothesis(los_mod,"grpTreatment < 0")[["hypothesis"]] %>%
  dplyr::select(Post.Prob)
#crude and combine
crude_los <- tbl_summary(data=bowel,by="grp",include="los",
                         statistic=list(all_continuous() ~ "{median} ({N_nonmiss}, {p25}, {p75})"),
                         digits = all_continuous() ~ 0) %>%
  modify_header(stat_1 = "{level}",stat_2 = "{level}") %>%
  as.data.frame()
los_res <- merge(crude_los,losrr) %>% merge(losh)
saveRDS(los_res,here("Output","Bo_los_res.RDS"))

#PRIOR SENSITIVITY
# los_mod_sens <- brm(los ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
#                data=bowel_mod,family=poisson,iter=10000,seed=46137112,
#                prior = c(prior(normal(0,10),class=b),
#                          prior(student_t(3,0,100),class=Intercept)))
#pr(sup)=0.22, same as other prior
#1.07, 0.89, 1.28 - pretty much same as other prior
```

## Device
```{r}
#read in the data
bowel_imp <- readRDS(here("Data","bowel_imp.RDS"))
#run the brms model
dev_mod <- brm_multiple(device ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
               data=bowel_imp,family=bernoulli,iter=15000,seed=46137112,
               prior = c(prior(normal(0,1000),class=b),
                         prior(student_t(3,0,100),class=Intercept)))
#check convergence of each imputation
dev_convergence <- convergence(chains=4,imps=20,mod=dev_mod)

#model results
dev_ci <- bayestestR::hdi(dev_mod)[2,3:4]
devh <- hypothesis(dev_mod,"grpTreatment > 0")[["hypothesis"]] %>%
  dplyr::select(Estimate,Post.Prob)
devor <- merge(dev_ci,devh) %>%
  mutate(OR = paste0(sprintf("%.2f",round(exp(Estimate),2))," (", round(exp(CI_low),2),", ",round(exp(CI_high),2),")"))
#crude and combine
crude_dev <- tbl_summary(data=bowel,by="grp",include="pt_device_hosp.factor",
                         statistic=list(all_dichotomous() ~ "{n}/{N} ({p}%)")) %>%
  modify_header(stat_1 = "{level}",stat_2 = "{level}") %>%
  as.data.frame()
dev_res <- merge(crude_dev,devor)

saveRDS(dev_res,here("Output","Bo_dev_res.RDS"))

#PRIOR SENSITIVITY - how different is it with prior of 10? (exp(10) is over 1000)
# dev_mod_sens <- brm_multiple(device ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
#                data=bowel_imp,family=bernoulli,iter=10000,seed=46137112,
#                prior = c(prior(normal(0,10),class=b),
#                          prior(student_t(3,0,100),class=Intercept)))
# dev_convergence_sens <- convergence(chains=4,imps=20,mod=dev_mod_sens)
#almost same results, very similar - 1.68 (95%CrI 0.066, 39.65) pr(sup)=0.62
```

## Meals
```{r}
#may want to dichotomise - the levels are tiny
bowel_imp <- readRDS(here("Data","bowel_imp.RDS"))

meal_mod <- brm_multiple(meals_bin ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
               data=bowel_imp,family=bernoulli,iter=10000,seed=46137112,
               prior = c(prior(normal(0,1000),class=b),
                         prior(student_t(3,0,100),class=Intercept)))
meal_convergence <- convergence(chains=4,imps=20,mod=meal_mod)

#model results
meal_ci <- bayestestR::hdi(meal_mod)[2,3:4]
mealh <- hypothesis(meal_mod,"grpTreatment > 0")[["hypothesis"]] %>%
  dplyr::select(Estimate,Post.Prob)
mealor <- merge(meal_ci,mealh) %>%
  mutate(OR = paste0(sprintf("%.2f",round(exp(Estimate),2))," (", round(exp(CI_low),2),", ",round(exp(CI_high),2),")"))
#crude and combine
crude_meals <- tbl_summary(data=bowel,by="grp",include="meals_bin",
                           statistic = list(all_dichotomous() ~ "{n}/{N} ({p}%)")) %>%
  modify_header(stat_1 = "{level}",stat_2 = "{level}") %>%
  as.data.frame()
meals_res <- merge(crude_meals,mealor)
saveRDS(meals_res,here("Output","Bo_meals_res.RDS"))

#PRIOR SENSITIVITY
# meal_mod_sens <- brm_multiple(meals_bin ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
#                data=bowel_imp,family=bernoulli,iter=10000,seed=46137112,
#                prior = c(prior(normal(0,10),class=b),
#                          prior(student_t(3,0,100),class=Intercept)))
#results pretty much the same
# 


```


## Snacks
```{r}
#suggest dichotomise into "any" = 1,2,3 vs "none" = 0
bowel_imp <- readRDS(here("Data","bowel_imp.RDS"))

snack_mod <- brm_multiple(snacks_bin ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
               data=bowel_imp,family=bernoulli,iter=10000,seed=46137112,
               prior = c(prior(normal(0,1000),class=b),
                         prior(student_t(3,0,100),class=Intercept)))
snack_convergence <- convergence(chains=4,imps=20,mod=snack_mod)

#model results
snack_ci <- bayestestR::hdi(snack_mod)[2,3:4]
snackh <- hypothesis(snack_mod,"grpTreatment > 0")[["hypothesis"]] %>%
  dplyr::select(Estimate,Post.Prob)
snackor <- merge(snack_ci,snackh) %>%
  mutate(OR = paste0(sprintf("%.2f",round(exp(Estimate),2))," (", round(exp(CI_low),2),", ",round(exp(CI_high),2),")"))
#crude and combine
crude_snack <- tbl_summary(data=bowel,by="grp",include="snacks_bin",
                           statistic = list(all_dichotomous() ~ "{n}/{N} ({p}%)")) %>%
  modify_header(stat_1 = "{level}",stat_2 = "{level}") %>%
  as.data.frame()
snack_res <- merge(crude_snack,snackor)
saveRDS(snack_res,here("Output","Bo_snack_res.RDS"))

#PRIOR SENSITIVITY
# snack_mod_sens <- brm_multiple(snacks_bin ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
#                data=bowel_imp,family=bernoulli,iter=10000,seed=46137112,
#                prior = c(prior(normal(0,10),class=b),
#                          prior(student_t(3,0,100),class=Intercept)))

```


## Fluids
```{r}
#suggest dichotomise into 4+ cups vs <=3 cups
bowel_imp <- readRDS(here("Data","bowel_imp.RDS"))

fluid_mod <- brm_multiple(fluid_bin ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
               data=bowel_imp,family=bernoulli,iter=10000,seed=46137112,
               prior = c(prior(normal(0,1000),class=b),
                         prior(student_t(3,0,100),class=Intercept)))
fluid_convergence <- convergence(chains=4,imps=20,mod=fluid_mod)

#model results
fluid_ci <- bayestestR::hdi(fluid_mod)[2,3:4]
fluidh <- hypothesis(fluid_mod,"grpTreatment > 0")[["hypothesis"]] %>%
  dplyr::select(Estimate,Post.Prob)
fluidor <- merge(fluid_ci,fluidh) %>%
  mutate(OR = paste0(sprintf("%.2f",round(exp(Estimate),2))," (", round(exp(CI_low),2),", ",round(exp(CI_high),2),")"))
#crude and combine
crude_fluid <- tbl_summary(data=bowel,by="grp",include="fluid_bin",
                           statistic = list(all_dichotomous() ~ "{n}/{N} ({p}%)")) %>%
  modify_header(stat_1 = "{level}",stat_2 = "{level}") %>%
  as.data.frame()
fluid_res <- merge(crude_fluid,fluidor)
saveRDS(fluid_res,here("Output","Bo_fluid_res.RDS"))

#PRIOR SENSITIVITY
# fluid_mod_sens <- brm_multiple(fluid_bin ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
#                data=bowel_imp,family=bernoulli,iter=10000,seed=46137112,
#                prior = c(prior(normal(0,10),class=b),
#                          prior(student_t(3,0,100),class=Intercept)))
#almost the same
```

## Time out of bed
```{r}
bowel_imp <- readRDS(here("Data","bowel_imp.RDS"))

oob_mod <- brm_multiple(bed_hosp_mins_total ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
               data=bowel_imp,family=poisson,iter=10000,seed=46137112,
               prior = c(prior(normal(0,1000),class=b),
                         prior(student_t(3,0,100),class=Intercept)))
oob_convergence <- convergence(chains=4,imps=20,mod=oob_mod)


#model results
oob_ci <- bayestestR::hdi(oob_mod)[2,3:4]
oobh <- hypothesis(oob_mod,"grpTreatment > 0")[["hypothesis"]] %>%
  dplyr::select(Estimate,Post.Prob)
oobrr <- merge(oob_ci,oobh) %>%
  mutate(RR = paste0(sprintf("%.2f",round(exp(Estimate),2))," (", round(exp(CI_low),2),", ",round(exp(CI_high),2),")"))
#crude and combine
crude_oob <- tbl_summary(data=bowel,by="grp",include="bed_hosp_mins_total",
                         statistic = list(all_continuous() ~ "{median} ({N_nonmiss}, {p25}, {p75})")) %>%
  modify_header(stat_1 = "{level}",stat_2 = "{level}") %>%
  as.data.frame()
oob_res <- merge(crude_oob,oobrr)
saveRDS(oob_res,here("Output","Bo_oob_res.RDS"))

#PRIOR SENSITIVITY
# oob_mod_sens <- brm_multiple(bed_hosp_mins_total ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
#                data=bowel_imp,family=poisson,iter=10000,seed=46137112,
#                prior = c(prior(normal(0,10),class=b),
#                          prior(student_t(3,0,100),class=Intercept)))
#quite similar results
```

## Time spent moving
```{r}
bowel_imp <- readRDS(here("Data","bowel_imp.RDS"))

move_mod <- brm_multiple(walk_hosp_mins_total ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
               data=bowel_imp,family=poisson,iter=10000,seed=46137112,
               prior = c(prior(normal(0,1000),class=b),
                         prior(student_t(3,0,100),class=Intercept)))
move_convergence <- convergence(chains=4,imps=20,mod=move_mod)

#model results
move_ci <- bayestestR::hdi(move_mod)[2,3:4]
moveh <- hypothesis(move_mod,"grpTreatment > 0")[["hypothesis"]] %>%
  dplyr::select(Estimate,Post.Prob)
moverr <- merge(move_ci,moveh) %>%
  mutate(RR = paste0(sprintf("%.2f",round(exp(Estimate),2))," (", round(exp(CI_low),2),", ",round(exp(CI_high),2),")"))
#crude and combine
crude_move <- tbl_summary(data=bowel,by="grp",include="walk_hosp_mins_total",
                         statistic = list(all_continuous() ~ "{median} ({N_nonmiss}, {p25}, {p75})")) %>%
  modify_header(stat_1 = "{level}",stat_2 = "{level}") %>%
  as.data.frame()
move_res <- merge(crude_move,moverr)
saveRDS(move_res,here("Output","Bo_move_res.RDS"))

#PRIOR SENSITIVITY
# move_mod_sens <- brm_multiple(walk_hosp_mins_total ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
#                data=bowel_imp,family=poisson,iter=10000,seed=46137112,
#                prior = c(prior(normal(0,10),class=b),
#                          prior(student_t(3,0,100),class=Intercept)))
#very similar results
```


## Breathing exercises
```{r}
bowel_imp <- readRDS(here("Data","bowel_imp.RDS"))

breath_mod <- brm_multiple(breath_bin ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
               data=bowel_imp,family=bernoulli,iter=10000,seed=46137112,
               prior = c(prior(normal(0,1000),class=b),
                         prior(student_t(3,0,100),class=Intercept)))
breath_convergence <- convergence(chains=4,imps=20,mod=breath_mod)

#model results
breath_ci <- bayestestR::hdi(breath_mod)[2,3:4]
breathh <- hypothesis(breath_mod,"grpTreatment > 0")[["hypothesis"]] %>%
  dplyr::select(Estimate,Post.Prob)
breathor <- merge(breath_ci,breathh) %>%
  mutate(OR = paste0(sprintf("%.2f",round(exp(Estimate),2))," (", round(exp(CI_low),2),", ",round(exp(CI_high),2),")"))
#crude and combine
crude_breath <- tbl_summary(data=bowel,by="grp",include="breath_bin",
                         statistic = list(all_dichotomous() ~ "{n}/{N} ({p}%)")) %>%
  modify_header(stat_1 = "{level}",stat_2 = "{level}") %>%
  as.data.frame()
breath_res <- merge(crude_breath,breathor)
saveRDS(breath_res,here("Output","Bo_breath_res.RDS"))

#PRIOR SENSITIVITY
# breath_mod_sens <- brm_multiple(breath_bin ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
#                data=bowel_imp,family=bernoulli,iter=10000,seed=46137112,
#                prior = c(prior(normal(0,10),class=b),
#                          prior(student_t(3,0,100),class=Intercept)))
#very similar results
```


## Knowledge
### Able to enhance own recovery
```{r}
bowel_imp <- readRDS(here("Data","bowel_imp.RDS"))

able_mod <- brm_multiple(able_bin ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
               data=bowel_imp,family=bernoulli,iter=15000,seed=46137112,
               prior = c(prior(normal(0,1000),class=b),
                         prior(student_t(3,0,100),class=Intercept)))
able_convergence <- convergence(chains=4,imps=20,mod=able_mod)

#model results
able_ci <- bayestestR::hdi(able_mod)[2,3:4]
ableh <- hypothesis(able_mod,"grpTreatment > 0")[["hypothesis"]] %>%
  dplyr::select(Estimate,Post.Prob)
ableor <- merge(able_ci,ableh) %>%
  mutate(OR = paste0(sprintf("%.2f",round(exp(Estimate),2))," (", round(exp(CI_low),2),", ",round(exp(CI_high),2),")"))
#crude and combine
crude_able <- tbl_summary(data=bowel,by="grp",include="able_bin",
                         statistic = list(all_dichotomous() ~ "{n}/{N} ({p}%)")) %>%
  modify_header(stat_1 = "{level}",stat_2 = "{level}") %>%
  as.data.frame()
able_res <- merge(crude_able,ableor)
saveRDS(able_res,here("Output","Bo_able_res.RDS"))

#PRIOR SENSITIVITY
# able_mod_sens <- brm_multiple(able_bin ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
#                data=bowel_imp,family=bernoulli,iter=15000,seed=46137112,
#                prior = c(prior(normal(0,10),class=b),
#                          prior(student_t(3,0,100),class=Intercept)))
#very similar results
```

### How to enhance own recovery
```{r}
bowel_imp <- readRDS(here("Data","bowel_imp.RDS"))

under_mod <- brm_multiple(understand_bin ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
               data=bowel_imp,family=bernoulli,iter=15000,seed=46137112,
               prior = c(prior(normal(0,1000),class=b),
                         prior(student_t(3,0,10),class=Intercept)))
under_convergence <- convergence(chains=4,imps=20,mod=under_mod)

#model results
under_ci <- bayestestR::hdi(under_mod)[2,3:4]
underh <- hypothesis(under_mod,"grpTreatment > 0")[["hypothesis"]] %>%
  dplyr::select(Estimate,Post.Prob)
underor <- merge(under_ci,underh) %>%
  mutate(OR = paste0(sprintf("%.2f",round(exp(Estimate),2))," (", round(exp(CI_low),2),", ",round(exp(CI_high),2),")"))
#crude and combine
crude_under <- tbl_summary(data=bowel,by="grp",include="understand_bin",
                         statistic = list(all_dichotomous() ~ "{n}/{N} ({p}%)")) %>%
  modify_header(stat_1 = "{level}",stat_2 = "{level}") %>%
  as.data.frame()
under_res <- merge(crude_under,underor)
saveRDS(under_res,here("Output","Bo_under_res.RDS"))

#PRIOR SENSITIVITY
# under_mod_sens <- brm_multiple(understand_bin ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
#                data=bowel_imp,family=bernoulli,iter=15000,seed=46137112,
#                prior = c(prior(normal(0,10),class=b),
#                          prior(student_t(3,0,10),class=Intercept)))
#very similar results

```

## Quality of recovery
```{r}
bowel_imp <- readRDS(here("Data","bowel_imp.RDS"))

qor_mod <- brm_multiple(qor_tot ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
               data=bowel_imp,family=gaussian,iter=10000,seed=46137112,
               prior = c(prior(normal(0,1000),class=b),
                         prior(student_t(3,0,100),class=Intercept)))
qor_convergence <- convergence(chains=4,imps=20,mod=qor_mod)

#model results
qor_ci <- bayestestR::hdi(qor_mod)[2,3:4]
qorh <- hypothesis(qor_mod,"grpTreatment > 0")[["hypothesis"]] %>%
  dplyr::select(Estimate,Post.Prob)
qormd <- merge(qor_ci,qorh) %>%
  mutate(MD = paste0(sprintf("%.2f",round(Estimate,2))," (", round(CI_low,2),", ",round(CI_high,2),")"))
#crude and combine
crude_qor <- tbl_summary(data=bowel,by="grp",include="qor_tot",
                         statistic = list(all_continuous() ~ "{mean} ({N_nonmiss}, {sd})"),
                         digits = all_continuous() ~ list(mean = 2, sd = 2, N_nonmiss = 0)) %>%
  modify_header(stat_1 = "{level}",stat_2 = "{level}") %>%
  as.data.frame()
qor_res <- merge(crude_qor,qormd)
saveRDS(qor_res,here("Output","Bo_qor_res.RDS"))

#PRIOR SENSITIVITY
# qor_mod_sens <- brm_multiple(qor_tot ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
#                data=bowel_imp,family=gaussian,iter=10000,seed=46137112,
#                prior = c(prior(normal(0,10),class=b),
#                          prior(student_t(3,0,100),class=Intercept)))
#Pr(sup) doesnt change, est changes but the main one is less informative so preferred I think
#This is a very conservative prior

```

## Combine all the models

```{r}

res_models <- bind_rows(los_res,dev_res,meals_res,snack_res,fluid_res,
                        oob_res,move_res,breath_res,able_res,under_res,qor_res) %>%
  filter(`**Characteristic**` != "Unknown") %>%
  rename(char = `**Characteristic**`) %>%
  mutate(est = case_when(!is.na(RR) ~ paste0("RR = ",RR),
                         !is.na(OR) ~ paste0("OR = ",OR),
                         !is.na(MD) ~ paste0("MD = ",MD)),
         outcome = case_when(char == "los" ~ "Length of stay (median (N, IQR))",
                             char == "pt_device_hosp.factor" ~ "Device (n/N %)",
                             char == "meals_bin" ~ "Meals (n/N %)",
                             char == "snacks_bin" ~ "Snacks (n/N %)",
                             char == "fluid_bin" ~ "Fluids (n/N %)",
                             char == "bed_hosp_mins_total" ~ "Time out of bed (median (N, IQR))",
                             char == "walk_hosp_mins_total" ~ "Time spent moving (median (N, IQR))",
                             char == "breath_bin" ~ "Breathing exercises (n/N %)",
                             char == "able_bin" ~ "Able to enhance own recovery (n/N %)",
                             char == "understand_bin" ~ "How to enhance own recovery (n/N %)",
                             char == "qor_tot" ~ "QOR at 2 days (mean (N, SD))")) %>%
  dplyr::select(outcome,A,B,est,Post.Prob)
saveRDS(res_models,here("Output","Bo_res_models.RDS"))
```

## Per protocol

### LOS

```{r}
bowel_mod_pp <- bowel %>%
  dplyr::select(grp,redcap_data_access_group,stoma.factor,
                surgery_type.factor,los,per_protocol)%>%
  mutate(grp = ifelse(grp == "A","Treatment","Control")) %>%
  filter(per_protocol == 1)

los_mod_pp <- brm(los ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
               data=bowel_mod_pp,family=poisson,iter=10000,seed=46137112,
               prior = c(prior(normal(0,1000),class=b),
                         prior(student_t(3,0,100),class=Intercept)))
#plot(los_mod_pp)
#pp_check(los_mod_pp)

#rate ratio
los_pprr <- avg_comparisons(los_mod_pp,variables=list(grp="pairwise"),
                         comparison = "ratio",ndraws=5000) %>%
  mutate(RR = paste0(sprintf("%.2f",round(estimate,2))," (", round(conf.low,2),", ",round(conf.high,2),")")) %>%
  dplyr::select(RR)
#p(sup)
los_pph <- hypothesis(los_mod_pp,"grpTreatment < 0")[["hypothesis"]] %>%
  dplyr::select(Post.Prob)
#crude and combine
crude_los_pp <- tbl_summary(data=bowel_mod_pp,by="grp",include="los",
                         statistic=list(all_continuous() ~ "{median} ({N_nonmiss}, {p25}, {p75})"),
                         digits = all_continuous() ~ 0) %>%
  modify_header(stat_1 = "{level}",stat_2 = "{level}") %>%
  as.data.frame()
los_res_pp <- merge(crude_los_pp,los_pprr) %>% merge(los_pph)
saveRDS(los_res_pp,here("Output","los_res_pp.RDS"))

#PRIOR SENSITIVITY
# los_mod_pp_sens <- brm(los ~ grp + redcap_data_access_group + stoma.factor + surgery_type.factor,
#                data=bowel_mod_pp,family=poisson,iter=10000,seed=46137112,
#                prior = c(prior(normal(0,10),class=b),
#                          prior(student_t(3,0,100),class=Intercept)))
#very similar results
```

#Second analysis phase

## Read in data
```{r}
bo_p2_imp <- readRDS(here("Data","p2bo_imp_long.RDS")) %>%
  mutate(treatment = ifelse(group_allocation == "1","Treatment","Control"),
         grp = ifelse(group_allocation == "1",1,0))

p2_bo_split <- split(x=bo_p2_imp,f=bo_p2_imp$.imp)

bo_p2 <- readRDS(here("Data","bo_p2.RDS")) %>%
  mutate(treatment = ifelse(group_allocation == "1","Treatment","Control"),
         grp = ifelse(group_allocation == "1",1,0))
  
```


## EORTC QLQ-C30  

At 30 and 90 days
Also has baseline value
```{r}
bo_p2_qlq <- bo_p2_imp %>%
  group_by(record_id,.imp) %>%
  mutate(summary_base = max(ifelse(time=="0",summary_score,NA),na.rm=T)) %>%
  filter(time %in% c(1,3))

qlq_split <- split(x=bo_p2_qlq,f=bo_p2_qlq$.imp)

c30_mod <- brm_multiple(summary_score ~ summary_base + grp*time + redcap_data_access_group.factor + stoma.factor + 
                        surgery_type.factor + (1|record_id),
                        data=qlq_split,family=gaussian,iter=12000,seed=46137112,
                        control = list(adapt_delta = 0.96),
                        prior = c(prior(normal(0,1000),class=b),
                        prior(student_t(3,0,100),class=Intercept),
                        prior(student_t(3,0,100),class=sigma)))
c30_convergence <- convergence(chains=4,imps=20,mod=c30_mod) %>% bind_rows()
#test it on just 1 imputation
#test <- loo(c30_mod)
#get the CrI and Pr(Sup) from the models
c30h <- hypothesis(c30_mod,c("grp > 0",
                             "grp + grp:time3 > 0"),
                     alpha=0.025)[["hypothesis"]] %>%
  dplyr::select(Hypothesis,Estimate,Post.Prob,CI.Lower,CI.Upper) %>%
  mutate(time = case_when(grepl("3",Hypothesis) ~ "3",
                          .default="1"),
         MD = paste0(sprintf("%.2f",round(Estimate,2))," (", round(CI.Lower,2),", ",round(CI.Upper,2),")"))

c30_res <- bo_p2 %>% 
  group_by(time,treatment) %>% 
  summarise(summary_score = paste0(round(mean(summary_score,na.rm=T),2)," (",sum(!is.na(summary_score)),
                                   ", ",round(sd(summary_score,na.rm=T),2),")")) %>% 
  pivot_wider(id_cols=time,names_from=treatment,values_from=summary_score) %>%
  filter(time != "hosp") %>%
  full_join(c30h,by="time")

saveRDS(c30_res,here("Output","bo_p2_c30.RDS"))
```

## Quality of Recovery

```{r}
bo_p2_qor <- bo_p2_imp %>%
  group_by(record_id,.imp) %>%
  mutate(qor_base = max(ifelse(time=="0",qor,NA),na.rm=T)) %>%
  filter(time %in% c("1","3","hosp"))
p2_bo_qor_split <- split(x=bo_p2_qor,f=bo_p2_qor$.imp)

qor_long_mod <- brm_multiple(qor ~ grp*time + qor_base + redcap_data_access_group.factor + 
                               stoma.factor + surgery_type.factor + (1|record_id),
               data=p2_bo_qor_split,family=gaussian,iter=26000,seed=46137112,
               prior = c(prior(normal(0,1000),class=b),
                         prior(student_t(3,0,100),class=Intercept),
                         prior(student_t(3,0,100),class=sigma)))
qor_long_convergence <- convergence(chains=4,imps=20,mod=qor_long_mod) %>% bind_rows()

#get the CrI and Pr(Sup) from the models
qorh <- hypothesis(qor_long_mod,c("grp > 0",
                                  "grp + grp:timehosp > 0",
                                  "grp + grp:time3 > 0"),
                     alpha=0.025)[["hypothesis"]] %>%
  dplyr::select(Hypothesis,Estimate,Post.Prob,CI.Lower,CI.Upper) %>%
  mutate(time = case_when(grepl("3",Hypothesis) ~ "3",
                          grepl("hosp",Hypothesis) ~ "hosp",
                          .default="1"),
         MD = paste0(sprintf("%.2f",round(Estimate,2))," (", round(CI.Lower,2),", ",round(CI.Upper,2),")"))

qor_res <- bo_p2 %>% 
  group_by(time,treatment) %>% 
  summarise(qor = paste0(round(mean(qor,na.rm=T),2)," (",sum(!is.na(qor)),
                                   ", ",round(sd(qor,na.rm=T),2),")")) %>% 
  pivot_wider(id_cols=time,names_from=treatment,values_from=qor) %>%
  #filter(time != "0") %>%
  full_join(qorh,by="time")

saveRDS(qor_res,here("Output","Bo_qor_res.RDS"))

```

## Days alive and out of hospital

Do these as separate models, otherwise the very high correlation between the outcomes might cause model issues.

### 30 days

```{r}
daoh_30 <- bo_p2 %>% 
  mutate(rev_daoh = 30 - daoh)

# daoh_imp <- readRDS(here("Data","p2bo_imp_long.RDS")) %>%
#   mutate(treatment = ifelse(group_allocation == "1","Treatment","Control"),
#          grp = ifelse(group_allocation == "1",1,0)) %>%
#   filter(.imp == 1) %>%
#   mutate(rev_daoh = 30 - daoh) %>%
#   group_by(record_id) %>%
#   mutate(summary_base = max(ifelse(time=="0",summary_score,NA),na.rm=T)) %>%
#   filter(time == "1")
  
daoh30_mod <- brm(rev_daoh ~ grp + redcap_data_access_group.factor + stoma.factor + surgery_type.factor,
                       data=daoh_30,family=poisson(),iter=10000,seed=46137112,
                       prior = c(prior(normal(0,10),class=b),
                                 #a prior for random intercept,
                                 prior(student_t(3,0,100),class=Intercept)))
#test <- loo(daoh30_mod)
#test_sum <- loo(daoh30_mods) #worse with eortc base, p_loo is 21 without, 23.8 with

#get the CrI and Pr(Sup) from the models
#flipping the outcome as we did the reverse
daoh30h <- hypothesis(daoh30_mod,c("grp < 0"))[["hypothesis"]] %>%
  dplyr::select(Post.Prob)

daohrr <- avg_comparisons(daoh30_mod,comparison = "ratio",variables = "grp") %>%
  mutate(estimate = 1/estimate,
         conf.low = 1/conf.low,
         conf.high = 1/conf.high,
         RR = paste0(round(estimate,2)," (",round(conf.high,2),", ",round(conf.low,2),")"),
         outcome = "DAOH") %>%
  cbind(daoh30h)
#combine with the crude outcome
daoh30_res <- bo_p2 %>% 
  filter(time == "1") %>%
  group_by(treatment) %>% 
  summarise(daoh = paste0(round(median(daoh,na.rm=T),2)," (",sum(!is.na(daoh)),
                                   ", ",round(quantile(daoh,probs=c(0.25),na.rm=T),2),
                          ", ",round(quantile(daoh,probs=c(0.75),na.rm=T),2),")")) %>% 
  pivot_wider(names_from=treatment,values_from=daoh) %>%
  cbind(daohrr)

saveRDS(daoh30_res,here("Output","bo_daoh30.RDS"))
```


### 90 days

```{r}
daoh_90 <- bo_p2 %>% filter(time == "3") %>% mutate(rev_daoh = 90 - daoh)

# daoh_imp <- readRDS(here("Data","p2bo_imp_long.RDS")) %>%
#   mutate(treatment = ifelse(group_allocation == "1","Treatment","Control"),
#          grp = ifelse(group_allocation == "1",1,0)) %>%
#   filter(.imp == 1) %>%
#   mutate(rev_daoh = 90 - daoh) %>%
#   group_by(record_id) %>%
#   mutate(summary_base = max(ifelse(time=="0",summary_score,NA),na.rm=T)) %>%
#   filter(time == "3")

daoh90_mod <- brm(rev_daoh ~ grp + redcap_data_access_group.factor + stoma.factor + surgery_type.factor,
                       data=daoh_90,family=negbinomial(),iter=10000,seed=46137112,
                       prior = c(prior(normal(0,10),class=b),
                                 #a prior for random intercept,
                                 prior(student_t(3,0,100),class=Intercept)))

#test <- loo(daoh90_mod)
#tests <- loo(daoh90_mods) #fits worse with summary score p_loo 7.3 with 8.7 without

#get the CrI and Pr(Sup) from the models
#flipping the outcome as we did the reverse
daoh90h <- hypothesis(daoh90_mod,c("grp < 0"))[["hypothesis"]] %>%
  dplyr::select(Post.Prob)

daohrr <- avg_comparisons(daoh90_mod,comparison = "ratio",variables = "grp") %>%
  mutate(estimate = 1/estimate,
         conf.low = 1/conf.low,
         conf.high = 1/conf.high,
         RR = paste0(round(estimate,2)," (",round(conf.high,2),", ",round(conf.low,2),")"),
         outcome = "DAOH") %>%
  cbind(daoh90h)

#combine with the crude outcome
daoh90_res <- bo_p2 %>% 
  filter(time == "3") %>%
  group_by(treatment) %>% 
  summarise(daoh = paste0(round(median(daoh,na.rm=T),2)," (",sum(!is.na(daoh)),
                                   ", ",round(quantile(daoh,probs=0.25,na.rm=T),2),
                          ", ",round(quantile(daoh,probs=0.75,na.rm=T),2),")")) %>% 
  pivot_wider(names_from=treatment,values_from=daoh) %>%
  cbind(daohrr)

saveRDS(daoh90_res,here("Output","bo_daoh90.RDS"))
```

## Health service utilisation

```{r}
cost_imp <- bo_p2_imp %>% mutate(costi = round(cost,0)) %>%
    group_by(record_id) %>%
  mutate(summary_base = max(ifelse(time=="0",summary_score,NA),na.rm=T)) %>%
  filter(time == "3")

cost_split <- split(x=cost_imp,f=cost_imp$.imp)

cost_mod <- brm(costi ~ grp + redcap_data_access_group.factor + stoma.factor + surgery_type.factor,
               data=cost_split,family=negbinomial,iter=13000,seed=46137112,
               prior = c(prior(normal(0,10),class=b),
                         prior(student_t(3,0,100),class=Intercept)))

#test <- loo(cost_mod)
#tests <- loo(cost_mods) #summary base is worse fit - with = 6.1, without = 5.5

#get the CrI and Pr(Sup) from the models
costh <- hypothesis(cost_mod,c("grp < 0"),alpha=0.025)[["hypothesis"]] %>%
  dplyr::select(Hypothesis,Estimate,Post.Prob,CI.Lower,CI.Upper) %>%
  mutate(RR = paste0(sprintf("%.2f",round(exp(Estimate),2))," (", round(exp(CI.Lower),2),", ",round(exp(CI.Upper),2),")"))

#combine with the crude outcome
cost_res <- bo_p2 %>% 
  filter(time == "3") %>%
  group_by(treatment) %>% 
  summarise(cost = paste0(round(median(cost,na.rm=T),2)," (",sum(!is.na(cost)),
                                   ", ",round(quantile(cost,probs=0.25,na.rm=T),2),
                          ", ",round(quantile(cost,probs=0.75,na.rm=T),2),")")) %>% 
  pivot_wider(names_from=treatment,values_from=cost) %>%
  cbind(costh)

saveRDS(cost_res,here("Output","cost_res.RDS"))
```

## Combine

```{r}
c30_res <- readRDS(here("Output","bo_p2_c30.RDS")) %>% mutate(outcome = "EORTC QLQ-C30 (mean (N, SD))")
qor_res <- readRDS(here("Output","Bo_qor_res.RDS")) %>% mutate(outcome = "Quality of recovery (mean (N, SD))")
daoh30_res <- readRDS(here("Output","bo_daoh30.RDS")) %>% mutate(time = "1",outcome = "DAOH (median (N, IQR))")
daoh90_res <- readRDS(here("Output","bo_daoh90.RDS")) %>% mutate(time = "3",outcome = "DAOH (median (N, IQR))")
cost_res <- readRDS(here("Output","cost_res.RDS")) %>% mutate(outcome = "Health service utilisation (median (N, IQR))",time="3")

res_models <- bind_rows(c30_res,qor_res,daoh30_res,daoh90_res,cost_res) %>%
  mutate(time = case_when(time == "1" ~ "30 days",
                          time == "3" ~ "90 days",
                          time == "hosp" ~ "2 days",
                          time == "0" ~ "Baseline"),
         order = row_number(),
         order = case_when(time == "Baseline" & outcome == "Quality of recovery (mean (N, SD))" ~ 1.9,
                           time == "2 days" & outcome == "Quality of recovery (mean (N, SD))" ~ 1.95,
                           .default=order),
         estimate = case_when(!is.na(MD) ~ paste0("MD = ",MD),
                              !is.na(RR) ~ paste0("RR = ",RR))) %>%
  arrange(order)%>%
  select(outcome,time,Treatment,Control,estimate,Post.Prob)

saveRDS(res_models,here("Output","bo_p2_res.RDS"))
```


